### 2025-06-02 Text/Reports/qualification_2024/chapters/1_intro.tex
R1 update: revised manuscript and reviewer responses added

The technology of Synthetic Aperture Radar (SAR) imaging operates on the comprehensive aperture principle to produce high-resolution SAR images.
In contrast to conventional optical remote sensing images, SAR images remain unaffected by external factors, making them suitable for identifying ground targets in diverse weather conditions and expansive areas.  
Consequently, SAR imagery has become essential for environmental monitoring, disaster management, agriculture, topography mapping, geological exploration, maritime navigation, defense, and climatology~\citep{Moreira2013, iglesias2013atmospheric, Mu2019}.  
However, the effective use of SAR
data depends on a thorough understanding of its statistical properties
because it is corrupted by speckle. This noise-like interference effect
is inherent in SAR data due to the coherent nature of the imaging
process~\citep{Argenti2013}.

@@

In this context, we present a set of three novel test statistics that
aim to distinguish between homogeneous and non-homogeneous returns,
particularly between gamma and \(\mathcal{G}^0\) distributed data,
assuming the number of looks is known.
 We use properties such as entropy and coefficient of variation.

Entropy is a fundamental concept in information theory with far-reaching
applications in pattern recognition, statistical physics, image
processing, edge detection and SAR image
analysis~\citep{Presse2013,MohammadDjafari2015,Avval2021, Nascimento2014,Nascimento2019}.
\citet{Shannon1948} this concept for a random variable to measure information and uncertainty. 
Shannon entropy is a crucial descriptive parameter in statistics, especially for evaluating data
dispersion and performing tests for normality, exponentiality and
uniformity~\citep{Wieczorkowski1999,Zamanzade2012}. 
Entropy estimation is challenging, especially when the model is unknown. In these cases,
non-parametric methods are used. Spacing methods have been discussed as
a non-parametric approach in Refs.~\citep{AlizadehNoughabi2010,Subhash2021}. This strategy is flexible
and robust because it does not enforce a model or parametric
constraints.

The coefficient of variation (CV), introduced in 1896 by~\citet{Pearson1896}, is a relative dispersion measure widely used
in various fields of applied statistics, including sampling,
biostatistics, medical and biological research, climatology and other
fields~\citep{hendricks1936sampling,Tian2005,SubrahmanyaNairy2003,Chankham2024}.
It facilitates the comparison of variability between different
populations and is particularly valuable for relating variables with
different units. This is because when the primary purpose is to compare
the variations of several variables, the standard deviation can only
serve as an adequate measure of variation if all variables are expressed
in the same unit of measurement and have identical means. If these
conditions are not met, then the CV is the relative measure that is
usually used in real applications. The variable with the highest CV
value has the largest relative dispersion around the mean
value~\citep{Banik2011}. The coefficient of variation is the primary
measure of heterogeneity in SAR data~\citep{Ulaby1986,Touzi1988}. We
study two ways of estimating the coefficient of variation.

The other parameter we study is the Shannon entropy. Different roughness
levels materialized as models for SAR data, have different entropy
values, but this fundamental quantity can also be estimated in a
model-agnostic way. We exploit this property and design a
bootstrap-improved non-parametric estimator for the Shannon entropy.

We devise test statistics based on these three estimators: the classical
coefficient of variation, a robust version, and the Shannon entropy
estimator. We apply these test statistics to generate maps of evidence
of homogeneity that reveal different types of targets in the SAR data.
We show that our proposed method is superior to existing approaches with
simulated data and SAR images.

@@

% colocar para citar los capitulos
\hypersetup{linkcolor=blue}
This document is organized as follows: 
Chapter~\ref{chp:background} provides a background on remote sensing and SAR images. 
Chapter~\ref{chp:methods} discusses methodological aspects, including statistical modeling, entropy estimation for intensity SAR data, and hypothesis testing. 
Chapter~\ref{chp:results} presents the main results obtained in our study with both simulated and actual data. Finally,
In Chapter~\ref{chp:conclusions}, the conclusions and future work are presented.
%Chapter ~\ref{chp:models} explains ...
%------------------------------------
%	HAZARD: OVERVIEW
%------------------------------------

%\section{Computing platforms}
%This  project required a vast amount of computing, preprocessing, analysis and plotting.

%------------------------------------
%	TODO STUFF
%------------------------------------
% \section{Additional sources to integrate in this chapter}
% \begin{itemize}
% \item check all TODOs
% \item read again \url{http://ec.europa.eu/environment/water/flood_risk/flood_atlas/pdf/handbook_goodpractice.pdf}
% \item include citation EU Floods Directive (2007/60/EC) , \cite{Mysiak2013, EUFD2007}
% \item section to talk about uncertainties? \citet{alfieri2014} has a good section about it
% \item Alps are the water tower of the Po plain, and for this reason I might want to give them a closer look. I could cite one of the many kotlarski works, or e.g. \citet[][]{Gobiet2014}. View the 'Alps' category in my Mendeley, it contains about 30 works on the subject
% \end{itemize}

>>>>

The technology of Synthetic Aperture Radar (SAR) imaging operates on the comprehensive aperture principle to produce high-resolution SAR images.
In contrast to conventional optical remote sensing images, SAR images remain unaffected by external factors, making them suitable for identifying ground targets in diverse weather conditions and expansive areas~\citep{Mu2019}.  
Consequently, SAR imagery has become essential for environmental monitoring~\citep{Amitrano2021}, crop mapping~\citep{DingleRobertson2020}, ship target detection~\citep{Wang2019}, and climatology~\citep{Zhao2023}.  
However, the effective use of SAR
data depends on a thorough understanding of its statistical properties
because it is corrupted by speckle. This noise-like interference effect
is inherent in SAR data due to the coherent nature of the imaging
process~\citep{Argenti2013}.

@@

Entropy is a fundamental concept in information theory with far-reaching
applications in pattern recognition~\citep{Avval2021}, statistical physics~\citep{Presse2013}, image
processing~\citep{MohammadDjafari2015}, edge detection~\citep{Nascimento2014} and SAR image
analysis~\citep{Nascimento2019}.
\citet{Shannon1948} introduced this concept for a random variable to measure information and uncertainty. 
Shannon entropy is a crucial descriptive parameter in statistics, especially for evaluating data
dispersion and performing tests for normality, exponentiality and
uniformity~\citep{Wieczorkowski1999,Zamanzade2012}.

Entropy estimation is challenging, especially when the model is unknown. In these cases,
non-parametric methods are used. Spacing methods have been discussed as
a non-parametric approach in Refs.~\citep{AlizadehNoughabi2010,Subhash2021}. This strategy is flexible
and robust because it does not enforce a model or parametric
constraints. In this context, we introduce a novel approach: a bootstrap-improved non-parametric estimator for Shannon entropy.

The coefficient of variation (CV), introduced in 1896 by~\citet{Pearson1896}, is a relative dispersion measure widely used
in various fields of applied statistics, including sampling~\citep{hendricks1936sampling},
biostatistics~\citep{Tian2005}, medical and biological research~\citep{SubrahmanyaNairy2003}, climatology~\citep{Chankham2024} and other
fields. It facilitates the comparison of variability between different
populations and is particularly valuable for relating variables with
different units. The variable with the highest CV
value has the largest relative dispersion around the mean
value~\citep{Banik2011}. The coefficient of variation is the primary
measure of heterogeneity in SAR data~\citep{Ulaby1986,Touzi1988}. 
In our work, we employ both the classical CV and a robust alternative, which is based on the ratio of the mean absolute deviation from the median (MnAD) to the median.  These measures offer a more detailed view of data variability.

The expected value, a critical component in both CV and the Shannon entropy, influences their interpretation. 
In CV calculations, the expected value serves as a reference point for assessing data dispersion relative to its average value, while in the Shannon entropy, the expected value also plays an important role, particularly within parametric distributions such as the Gamma or the \({G}_I^0\)  distribution. 
In these cases, the mean appears in the analytical entropy formula, indicating that the entropy of the distribution is influenced by the mean value of the data.

When applying these statistical concepts to SAR image analysis, we aim to discern between homogeneous and non-homogeneous regions, contributing to the improved interpretation and understanding of SAR data.  In our study, we devise test statistics based on these three estimators: the classical coefficient of variation, a robust version, and the Shannon entropy estimator. We apply these test statistics to generate maps of evidence of homogeneity that reveal different types of targets in the SAR data, assuming the number of looks is known. Likewise, we show that our proposed method is superior to existing approaches with simulated data and SAR images.

@@

% colocar para citar los capitulos
\hypersetup{linkcolor=blue}
This document is organized as follows: 
Chapter~\ref{chp:background} provides a background on remote sensing and SAR images. 
Chapter~\ref{chp:methods} discusses methodological aspects, including statistical modeling, entropy estimation for intensity SAR data, and hypothesis testing. 
Chapter~\ref{chp:results} presents the main results obtained in our study with both simulated and actual data. Finally,
In Chapter~\ref{chp:conclusions}, the conclusions and future work are presented.

% \item section to talk about uncertainties? \citet{alfieri2014} has a good section about it
% \item Alps are the water tower of the Po plain, and for this reason I might want to give them a closer look. I could cite one of the many kotlarski works, or e.g. \citet[][]{Gobiet2014}. View the 'Alps' category in my Mendeley, it contains about 30 works on the subject
% \end{itemize}

### 2025-06-02 Text/Reports/qualification_2024/chapters/2_background.tex
R1 update: revised manuscript and reviewer responses added

There are two types of remote sensors acquiring information in fundamentally different ways.
Optical sensors are passive imaging devices, which means they take advantage of existing sources of illumination and the natural reflective properties of objects to form
images. 
In the case of optical Earth observation imagery, the sun is used as a global illumination source. 
Thus optical image sensors tend to operate on a sun-synchronous orbit to ensure the scene is well illuminated during acquisition. 
On the other hand, SAR uses the principle of active sensing, whereby the sensor acts as both an illumination source as well as the acquisition unit. 
A scene is thus imaged by alternating between emitting bursts of EM-radiation and measuring the strength and time delay of the reflected signal. 
The proportion of the signal which is reflected towards the sensor by an object is known as \textit{backscatter}.
\subsection{The Electromagnetic Spectrum}
Given that objects interact (reflect and absorb) different parts of the EM-spectrum in a unique manner, several realizations of optical and SAR sensors exist, each designed for
a specific purpose and to take advantage of the properties observable within a particular part of the EM-spectrum.
Optical sensors detect very high-frequency radiation within the visible to thermal infrared section of the EM-spectrum.
In contrast, SAR sensors operate at much lower frequencies within the microwave section of the spectrum.
It is this use of lower frequencies which allow SAR sensors to be mostly independent of atmospheric conditions.
The same, however, is not true for optical sensors which require clear atmospheric conditions to observe ground level reflections within the visible light and near-infrared spectrum.

@@

\setlength{\tabcolsep}{6.0pt}
\begin{table}[!htb]
\caption[Different bands in microwave remote sensing]{\label{tab:bands}Different bands in microwave remote sensing,~\citep{Moreira2013}.}
\centering
\begin{tabular}[t]{cccccc}
\toprule
 \textbf{Band} & \multicolumn{1}{c}{\textbf{Frequency} $\left[\text{GHz}\right]$} & \multicolumn{1}{c}{\textbf{Wavelength} $\left[\text{cm}\right]$} \\
\midrule
Ka & $\,25$--$40$ & $\phantom{-}0.75$--$1.2$ \\
Ku & $\,12$--$17.6$ & $\phantom{-}\,1.7$--$2.5$ \\
X  & $7.5$--$12$  & $2.5$--$4$   \\
C  & $ 3.75$--$7.5$ & $\phantom{-}4$--$8$   \\
S  & $2$--$3.75$   & $\phantom{-}\,8$--$15$    \\
L  & $1$--$2$   & $\phantom{-}15$--$30$   \\
P  & $0.3$--$1$ & $\phantom{-}\,30$--$100$  \\
\bottomrule
\end{tabular}
\end{table}
Table~\ref{tab:spacebornes} gives some examples of commonly used spaceborne SAR sensors, currently in operation. 
These include Radarsat-2 from the Canadian Space Agency (CSA), TerraSAR-X from the German Aerospace Center (DLR), ALOS PALSAR-2 from the Japan Aerospace Exploration Agency (JAXA), and Sentinel-1A/1B from the European Space Agency (ESA) (see Figure~\ref{fig:Sentinel-1n}), as well as Gaofen-3 from the China National Space Administration (CNSA).
\begin{figure}[H]
    \centering
    \includegraphics[width=0.6\textwidth]{../../Images/PNG/ESA_radar.png}
    %\decoRule
    \caption[Sentinel-1 radar.]{Sentinel-1 radar, (source \href{https://www.esa.int/Applications/Observing_the_Earth/Copernicus/Sentinel-1/Changing_lands}{ESA}).}
    \label{fig:Sentinel-1n}
\end{figure}

@@

In Figure~\ref{fig:speckle}, SAR images depict speckle, a characteristic of SAR data due to its coherent imaging process. 
The images distinguishes between fully developed speckle in homogeneous regions (Figure~\ref{fig:speckle} \ref{fig:s1}) and heterogeneous clutter in regions such as urban areas (Figure~\ref{fig:speckle} \ref{fig:s2}).
\begin{figure}[H]
  \centering
  \begin{subfigure}[b]{0.48\textwidth}
    \centering
    \includegraphics[width=.8\textwidth]{../../Images/PNG/Intensity_MG.png}
    \caption{Fully developed speckle (Homogeneous areas)}
    \label{fig:s1}
  \end{subfigure}
  \hfill
  \begin{subfigure}[b]{0.45\textwidth}
    \centering
    \includegraphics[width=.85\textwidth]{../../Images/PNG/urban_intensity.png}
    \caption{Heterogeneous clutter }
    \label{fig:s2}
  \end{subfigure}
  \caption{SAR image with speckle noise}
  \label{fig:speckle}
\end{figure}

>>>>

There are two types of remote sensors acquiring information in fundamentally different ways.
Optical sensors are passive imaging devices, which means they take advantage of existing sources of illumination and the natural reflective properties of objects to form
images. 
In the case of optical Earth observation imagery, the sun is used as a global illumination source. 
Thus, optical image sensors tend to operate on a sun-synchronous orbit to ensure the scene is well illuminated during acquisition. 
On the other hand, SAR uses the principle of active sensing, whereby the sensor acts as both an illumination source as well as the acquisition unit. 
A scene is thus imaged by alternating between emitting bursts of EM-radiation and measuring the strength and time delay of the reflected signal. 
The proportion of the signal which is reflected towards the sensor by an object is known as \textit{backscatter}.
\subsection{The Electromagnetic Spectrum}
Given that objects interact (reflect and absorb) different parts of the EM-spectrum in a unique manner, several realizations of optical and SAR sensors exist, each designed for
a specific purpose and to take advantage of the properties observable within a particular part of the EM-spectrum.
Optical sensors detect very high-frequency radiation within the visible to thermal infrared section of the EM-spectrum.
In contrast, SAR sensors operate at much lower frequencies within the microwave section of the spectrum.
It is this use of lower frequencies which allow SAR sensors to be mostly independent of atmospheric conditions.
The same, however, is not true for optical sensors which require clear atmospheric conditions to observe ground level reflections within the visible light and near-infrared spectrum.

@@

\setlength{\tabcolsep}{6.0pt}
\begin{table}[!htb]
\caption[Different bands in microwave remote sensing]{\label{tab:bands}Different bands in microwave remote sensing,~\citep{Moreira2013}.}
\centering
\begin{tabular}[t]{cccccc}
\toprule
 \textbf{Band} & \multicolumn{1}{c}{\textbf{Frequency} $\left[\text{GHz}\right]$} & \multicolumn{1}{c}{\textbf{Wavelength} $\left[\text{cm}\right]$} \\
\midrule
Ka & $\,25.0$--$40.0$ & $\phantom{-}0.75$--$1.2$ \\
Ku & $\,12.0$--$17.6$ & $\phantom{-}\,1.7$--$2.5$ \\
X  & $\phantom{-}7.5$--$12.0$  & $\phantom{-}2.5$--$4.0$   \\
C  & $ 3.7$--$7.5$ & $\phantom{-}4.0$--$8.0$   \\
S  & $2.0$--$3.7$   & $\phantom{-}\,8.0$--$15$    \\
L  & $1.0$--$2.0$   & $\phantom{-}15$--$30$   \\
P  & $0.3$--$1.0$ & $\phantom{-}\,30$--$100$  \\
\bottomrule
\end{tabular}
\end{table}

%\begin{longtable}{@{}
			%c
			%S[table-format=-4.1]@{--}S[table-format=2.5]
			%S[table-format=-2.1,
		%table-align-text-pre=false]@{--}S[table-format=2.6, table-align-text-pre=false]
			%%S[table-format=-2.1]@{--}S[table-format=2.3]
			%@{}}
		%%\caption{Las entradas de esta tabla están centradas correctamente, pero las distancias son difíciles de ajustar.}%
		%\caption[Different bands in microwave remote sensing]{\label{tab:bands}Different bands in microwave remote sensing,~\citep{Moreira2013}.}
		%\\
		%\toprule
		%\text{Band}            &
		%\multicolumn{2}{c}{\text{\textbf{Frequency} $\left[\text{GHz}\right]$}}        &
		%\multicolumn{2}{c}{\textbf{Wavelength} $\left[\text{cm}\right]$}          \\
		%\midrule
		%\endfirsthead
		%\midrule
		%\text{Band}            &
		%\multicolumn{2}{c}{\text{\textbf{Frequency} $\left[\text{GHz}\right]$}}        &
		%\multicolumn{2}{c}{\textbf{Wavelength} $\left[\text{cm}\right]$}          \\
		%\midrule
		%\endhead
		%Ka & 25.0 & 40.0 & 0.7 & 1.2 \\
		%Ku & 12.0 & 17.6 & 1.7 & 2.5 \\
		%X  & 7.5 & 12.0 & 2.5 & 4.0 \\
		%C  & 3.7 &7.5 & 4.0 & 8.0 \\
		%S  & 2.0 & 3.7 & 8.0 & 15 \\
		%L  & 1.0 & 2.0 & 15.0 & 30 \\
		%P  & 0.3 & 1.0 & 30.0 & 100 \\
		%\bottomrule
	%\end{longtable}

Table~\ref{tab:spacebornes} gives some examples of commonly used spaceborne SAR sensors, currently in operation. 
These include Radarsat-2 from the Canadian Space Agency (CSA), TerraSAR-X from the German Aerospace Center (DLR), ALOS PALSAR-2 from the Japan Aerospace Exploration Agency (JAXA), and Sentinel-1A/1B from the European Space Agency (ESA) (see Figure~\ref{fig:Sentinel-1n}), as well as Gaofen-3 from the China National Space Administration (CNSA).
\begin{figure}[H]
    \centering
    \includegraphics[width=0.6\textwidth]{../../Images/PNG/ESA_radar.png}
    %\decoRule
    \caption[Sentinel-1 radar.]{Sentinel-1 radar, (source \href{https://www.esa.int/Applications/Observing_the_Earth/Copernicus/Sentinel-1/Changing_lands}{ESA}).}
    \label{fig:Sentinel-1n}
\end{figure}

@@

In Figure~\ref{fig:speckle}, SAR images depict speckle, a characteristic of SAR data due to its coherent imaging process. 
The images distinguish between fully developed speckle in homogeneous regions (Figure~\ref{fig:speckle} \ref{fig:s1}) and heterogeneous clutter in regions such as urban areas (Figure~\ref{fig:speckle} \ref{fig:s2}).
\begin{figure}[H]
  \centering
  \begin{subfigure}[b]{0.48\textwidth}
    \centering
    \includegraphics[width=.8\textwidth]{../../Images/PNG/Intensity_MG.png}
    \caption{Fully developed speckle (Homogeneous areas)}
    \label{fig:s1}
  \end{subfigure}
  \hfill
  \begin{subfigure}[b]{0.45\textwidth}
    \centering
    \includegraphics[width=.85\textwidth]{../../Images/PNG/urban_intensity.png}
    \caption{Heterogeneous clutter }
    \label{fig:s2}
  \end{subfigure}
  \caption{SAR image with speckle noise}
  \label{fig:speckle}
\end{figure}

### 2025-06-02 Text/Reports/qualification_2024/chapters/3_methods.tex
R1 update: revised manuscript and reviewer responses added

The problem of the non-parametric estimation of \(H(Z)\) has been
studied by many authors,
including~\citep{vasicek1976test,correa1995new,Wieczorkowski1999,Yee2015,AlOmari2019}.
Their proposals use estimators based on differences between order
statistics: spacings.

@@

Several authors have explored adaptations  to Vasicek's estimator. We consider the following entropy estimators variants, as discussed by~\citep{AlOmari2016,Cassetti2022}.

\citet{Bert1992} proposed a new estimator of entropy given by:
 \begin{equation}
\label{E:VanEs}
	\widehat{H}_{\text{VE}}(\bm{Z})=\frac{1}{n-m}\sum_{i=1}^{n-m}\log\left[\frac{n+1}{m}\left(Z_{(i+m)}-Z_{(i)}\right)\right]	+\sum_{k=m}^n\frac{1}{k}+\log\frac{m}{n+1}.
\end{equation}
Under some conditions, Van Es proved asymptotic normality of this estimator.

\citet{Ebrahimi1994} adjusted the weights of Vasicek's estimator, in order to take into account the fact that the differences are truncated around the smallest and the largest data points.
 Specifically, $Z_{(i+m)}-Z_{(i-m)}$ is substituted with  $Z_{(i+m)}-Z_{(1)}$ when $i \leq m$ and $Z_{(i+m)}-Z_{(i-m)}$ is replaced by $Z_{(n)}-Z_{(i-m)}$ when $i \geq n-m+1$. Their estimator is given by:
  \begin{equation}
	\label{E:Ebrahimi}
  \widehat{H}_{\text{E}}(\bm{Z})=\frac{1}{n} \sum_{i=1}^n \log \left[\frac{n}{c_i m}\left(Z_{(i+m)}-Z_{(i-m)}\right)\right],
  \end{equation} where \[
  c_i=\begin{cases}1+(i-1) / m & \text { if } \quad 1 \leq i \leq m, \\ 2 & \text { if }\quad m+1 \leq i \leq n-m,\\ 1+(n-i) / m & \text { if }\quad n-m+1 \leq i \leq n.\end{cases}
  \]

 \citet{correa1995new} suggested another modification of Vasicek's estimator. 
In estimation the density $f$ of $F$ in the interval $\left(Z_{(i-m)}, Z_{(i+m)}\right)$ he used a local linear model based on $2 m+1$ points: $F\left(Z_{(j)}\right)=a+b Z_{(j)}+\varepsilon, j=m-i, \ldots, m+i$. This yields a following estimator:
  \begin{equation}
	\label{E:Correa}
  \widehat{H}_{\text{C}}(\bm{Z})=-\frac{1}{n} \sum_{i=1}^n \log \frac{\sum_{j=i-m}^{i+m}(j-i)\left(Z_{(j)}-\overline{Z}_{(i)}\right)}{n\sum_{j=i-m}^{i+m}\left(Z_{(j)}-\overline{Z}_{(i)}\right)^2},
  \end{equation} where
  \(\widebar{Z}_{(i)}=(2 m+1)^{-1} \sum_{j=i-m}^{i+m} Z_{(j)}\),
  \(m< \frac{n}{2}\), \(Z_{(i)}=Z_{(1)}\) for \(i<1\) and
  \(Z_{(i)}=Z_{(n)}\) for \(i>n\). Based on simulations, he showed that
  his estimator has a smaller mean square error than Vasicek's approach.

\citet{rohtua} modify the coefficients of~\eqref{E:Ebrahimi} as: 
\begin{equation}
\label{E:NA}
\widehat{H}_{\text{NA}}(\bm{Z})=\frac{1}{n} \sum_{i=1}^n \log \left[\frac{n}{a_i m}\left(Z_{(i+m)}-Z_{(i-m)}\right)\right],
\end{equation}
where
$$
a_i=\begin{cases}
1 & \text { if } \quad 1 \leq i \leq m, \\
2 & \text { if } \quad m+1 \leq i \leq n-m, \\
1 & \text { if } \quad n-m+1 \leq i \leq n,
\end{cases}
$$
and $Z_{(i-m)}=Z_{(1)}$ for $i \leq m$ and $Z_{(i+m)}=Z_{(n)}$ for $i \geq n-m$.

\citet{IbrahimAlOmari2014} suggested the following estimator:
	\begin{equation}
\label{E:AO}
  \widehat{H}_{\text{AO}}(\bm{Z})=\frac{1}{n} \sum_{i=1}^n \log \left[\frac{n}{\omega_i m}\left(Z_{(i+m)}-Z_{(i-m)}\right)\right],
 \end{equation}
	where \[
  \omega_i= \begin{cases}3/2 & \text { if }\quad 1 \leq i \leq m, \\ 2 & \text { if }\quad m+1 \leq i \leq n-m, \\ 3/2 & \text { if } \quad n-m+1 \leq i \leq n,\end{cases}
  \] in which \(Z_{(i-m)}=Z_{(1)}\) for \(i \leq m\), and
  \(Z_{(i+m)}=Z_{(n)}\) for \(i \geq n-m\).

@@

Bootstrap is essentially a resampling algorithm introduced by \citep{Efron1979}. 
It offers a simulation-based approach for estimating various statistics of interest for a random variable.
It can be used in both parametric and nonparametric settings. 
The main idea  consists of creating new datasets from an existing one by resampling with repetition~\citep{Michelucci2021}.

@@

Our objective is to devise a unbiased estimator. To achieve this, we introduce an "ideal estimator" \(\widecheck{\theta}(\bm{Z})\) using the bias information from~\eqref{Eq:bias1}:
\begin{align}
\label{Eq:bias2}
\widecheck{\theta}(\bm{Z}) &= \widehat{\theta}(\bm{Z}) - \bias\left(\widehat{\theta}(\bm{Z})\right),\nonumber\\
&=\widehat{\theta}(\bm{Z})-\left[\mathbb{E}\big[\widehat{\theta}(\bm{Z})\big] - \theta\right],\nonumber\\
&=\widehat{\theta}(\bm{Z})+\theta-\mathbb{E}\big[\widehat{\theta}(\bm{Z})\big].
\end{align}
However, \(\widecheck{\theta}(\bm{Z})\) is not yet an estimator, because it depend on the true parameter, but \(\theta\) can be calculated using a non-parametric approach, and the average of individual estimators can be obtained using the bootstrap technique. This leads to the formulation of a new unbiased estimator, denoted as \(\widetilde{H}\), from ~\eqref{Eq:bias2} we have:
\begin{align}
\label{Eq:bias3}
\widetilde{H}%&= \widehat{\theta}(\bm{Z})+\widehat{\theta}(\bm{Z})- \frac{1}{B}\sum_{b=1}^B \widehat{\theta}_b(\bm{Z}^{(b)}),\nonumber\\
= 2\widehat{\theta}(\bm{Z}) - \frac{1}{B}\sum_{b=1}^B \widehat{\theta}_b(\bm{Z}^{(b)}),
\end{align} where \(B\) the number of bootstrap replications. For each iteration \(b\) (from \(1\) to \(B\)), a sample \(\bm{Z}^{(b)}\) is generated, and the corresponding bootstrap estimator \(\widehat{\theta}_b(\bm{Z}^{(b)})\) is calculated. 
This methodology significantly reduces bias, resulting in more accurate and robust entropy estimations. The original estimators of Vasicek~\eqref{E:Vas}, Van Es~\eqref{E:VanEs}, Ebrahimi~\eqref{E:Ebrahimi}, Correa~\eqref{E:Correa}, Noughabi~\eqref{E:NA}, and Al-Omari~\eqref{E:AO}  are now referred to as the proposed bootstrap-improved versions:
\(\widetilde{H}_{\text{V}}\),
\(\widetilde{H}_{\text{VE}}\), 
\(\widetilde{H}_{\text{E}}\),
\(\widetilde{H}_{\text{C}}\),
\(\widetilde{H}_{\text{NA}}\), and 
\(\widetilde{H}_{\text{AO}}\),
respectively.

>>>>

The problem of the non-parametric estimation of \(H(Z)\) has been
studied by many authors, including~\citet{vasicek1976test},~\citet{correa1995new},~\citet{Wieczorkowski1999},~\citet{Yee2015}, and~\citet{AlOmari2019}.
Their proposals use estimators based on differences between order statistics, known as spacings.

@@

Several authors have explored adaptations  to Vasicek's estimator. We consider the following entropy estimators variants, as discussed by~\citet{Cassetti2022}.

\citet{Bert1992} proposed a new estimator of entropy given by:
 \begin{equation}
\label{E:VanEs}
	\widehat{H}_{\text{VE}}(\bm{Z})=\frac{1}{n-m}\sum_{i=1}^{n-m}\ln\left[\frac{n+1}{m}\left(Z_{(i+m)}-Z_{(i)}\right)\right]	+\sum_{k=m}^n\frac{1}{k}+\ln\frac{m}{n+1}.
\end{equation}
Under some conditions, Van Es proved asymptotic normality of this estimator.

\citet{Ebrahimi1994} adjusted the weights of Vasicek's estimator, in order to take into account the fact that the differences are truncated around the smallest and the largest data points.
 Specifically, $Z_{(i+m)}-Z_{(i-m)}$ is substituted with  $Z_{(i+m)}-Z_{(1)}$ when $i \leq m$ and $Z_{(i+m)}-Z_{(i-m)}$ is replaced by $Z_{(n)}-Z_{(i-m)}$ when $i \geq n-m+1$. Their estimator is given by:
  \begin{equation}
	\label{E:Ebrahimi}
  \widehat{H}_{\text{E}}(\bm{Z})=\frac{1}{n} \sum_{i=1}^n \ln \left[\frac{n}{c_i m}\left(Z_{(i+m)}-Z_{(i-m)}\right)\right],
  \end{equation} where \[
  c_i=\begin{cases}1+(i-1) / m & \text { if } \quad 1 \leq i \leq m, \\ 2 & \text { if }\quad m+1 \leq i \leq n-m,\\ 1+(n-i) / m & \text { if }\quad n-m+1 \leq i \leq n.\end{cases}
  \]

 \citet{correa1995new} suggested another modification of Vasicek's estimator. 
In estimation the density $f$ of $F$ in the interval $\left(Z_{(i-m)}, Z_{(i+m)}\right)$ he used a local linear model based on $2 m+1$ points: $F\left(Z_{(j)}\right)=a+b Z_{(j)}+\varepsilon, j=m-i, \ldots, m+i$. This yields a following estimator:
  \begin{equation}
	\label{E:Correa}
  \widehat{H}_{\text{C}}(\bm{Z})=-\frac{1}{n} \sum_{i=1}^n \ln \frac{\sum_{j=i-m}^{i+m}(j-i)\left(Z_{(j)}-\overline{Z}_{(i)}\right)}{n\sum_{j=i-m}^{i+m}\left(Z_{(j)}-\overline{Z}_{(i)}\right)^2},
  \end{equation} where
  \(\widebar{Z}_{(i)}=(2 m+1)^{-1} \sum_{j=i-m}^{i+m} Z_{(j)}\),
  \(m< \frac{n}{2}\), \(Z_{(i)}=Z_{(1)}\) for \(i<1\) and
  \(Z_{(i)}=Z_{(n)}\) for \(i>n\). Based on simulations, he showed that
  his estimator has a smaller mean square error than Vasicek's approach.

\citet{rohtua} modify the coefficients of~\eqref{E:Ebrahimi} as: 
\begin{equation}
\label{E:NA}
\widehat{H}_{\text{NA}}(\bm{Z})=\frac{1}{n} \sum_{i=1}^n \ln \left[\frac{n}{a_i m}\left(Z_{(i+m)}-Z_{(i-m)}\right)\right],
\end{equation}
where
$$
a_i=\begin{cases}
1 & \text { if } \quad 1 \leq i \leq m, \\
2 & \text { if } \quad m+1 \leq i \leq n-m, \\
1 & \text { if } \quad n-m+1 \leq i \leq n,
\end{cases}
$$
and $Z_{(i-m)}=Z_{(1)}$ for $i \leq m$ and $Z_{(i+m)}=Z_{(n)}$ for $i \geq n-m$.

\citet{IbrahimAlOmari2014} suggested the following estimator:
	\begin{equation}
\label{E:AO}
  \widehat{H}_{\text{AO}}(\bm{Z})=\frac{1}{n} \sum_{i=1}^n \ln \left[\frac{n}{\omega_i m}\left(Z_{(i+m)}-Z_{(i-m)}\right)\right],
 \end{equation}
	where \[
  \omega_i= \begin{cases}3/2 & \text { if }\quad 1 \leq i \leq m, \\ 2 & \text { if }\quad m+1 \leq i \leq n-m, \\ 3/2 & \text { if } \quad n-m+1 \leq i \leq n,\end{cases}
  \] in which \(Z_{(i-m)}=Z_{(1)}\) for \(i \leq m\), and
  \(Z_{(i+m)}=Z_{(n)}\) for \(i \geq n-m\).

@@

Bootstrap is essentially a resampling algorithm introduced by \citep{Efron1979}. 
It offers a simulation-based approach for estimating various statistics of interest for a random variable.
It can be used in both parametric and non-parametric settings. 
The main idea  consists of creating new datasets from an existing one by resampling with repetition~\citep{Michelucci2021}.

@@

Our objective is to devise an unbiased estimator. To achieve this, we introduce an "ideal estimator" \(\widecheck{\theta}(\bm{Z})\) using the bias information from~\eqref{Eq:bias1}:
\begin{align}
\label{Eq:bias2}
\widecheck{\theta}(\bm{Z}) &= \widehat{\theta}(\bm{Z}) - \bias\left(\widehat{\theta}(\bm{Z})\right),\nonumber\\
&=\widehat{\theta}(\bm{Z})-\left[\mathbb{E}\big[\widehat{\theta}(\bm{Z})\big] - \theta\right],\nonumber\\
&=\widehat{\theta}(\bm{Z})+\theta-\mathbb{E}\big[\widehat{\theta}(\bm{Z})\big].
\end{align}
However, \(\widecheck{\theta}(\bm{Z})\) is not yet an estimator, because it depends on the true parameter, but \(\theta\) can be calculated using a non-parametric approach, and the average of individual estimators can be obtained using the bootstrap technique. This leads to the formulation of a new unbiased estimator, denoted as \(\widetilde{H}\), from ~\eqref{Eq:bias2} we have:
\begin{align}
\label{Eq:bias3}
\widetilde{H}%&= \widehat{\theta}(\bm{Z})+\widehat{\theta}(\bm{Z})- \frac{1}{B}\sum_{b=1}^B \widehat{\theta}_b(\bm{Z}^{(b)}),\nonumber\\
= 2\widehat{\theta}(\bm{Z}) - \frac{1}{B}\sum_{b=1}^B \widehat{\theta}_b(\bm{Z}^{(b)}),
\end{align} where \(B\) the number of bootstrap replications. For each iteration \(b\) (from \(1\) to \(B\)), a sample \(\bm{Z}^{(b)}\) is generated, and the corresponding bootstrap estimator \(\widehat{\theta}_b(\bm{Z}^{(b)})\) is calculated. 
This methodology significantly reduces bias, resulting in more accurate and robust entropy estimations. The original estimators of Vasicek~\eqref{E:Vas}, Van Es~\eqref{E:VanEs}, Ebrahimi~\eqref{E:Ebrahimi}, Correa~\eqref{E:Correa}, Noughabi~\eqref{E:NA}, and Al-Omari~\eqref{E:AO}  are now referred to as the proposed bootstrap-improved versions:
\(\widetilde{H}_{\text{V}}\),
\(\widetilde{H}_{\text{VE}}\), 
\(\widetilde{H}_{\text{E}}\),
\(\widetilde{H}_{\text{C}}\),
\(\widetilde{H}_{\text{NA}}\), and 
\(\widetilde{H}_{\text{AO}}\),
respectively.
